---
title: "Data Science for Business - Project"
author: "Broersen Doortje, Prickaerts Charles, Van Rijn Rens, Feutray Charles, Ammann David"
output:
  html_document:
    css: ~/00_INSEAD/03_P3/DSfB/Projects/INSEADAnalytics/AnalyticsStyles/default.css
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    includes:
      in_header: ~/00_INSEAD/03_P3/DSfB/Projects/INSEADAnalytics/AnalyticsStyles/default.sty
always_allow_html: yes
---

```{r echo=FALSE, message=FALSE}
make_pdf_file = 0 # SET THIS TO 1 IF WE COMPILE PDF FILE, 0 OTHERWISE (FOR HTML)

source("~/00_INSEAD/03_P3/DSfB/Projects/INSEADAnalytics/AnalyticsLibraries/library.R")
source("~/00_INSEAD/03_P3/DSfB/Projects/INSEADAnalytics/AnalyticsLibraries/heatmapOutput.R")

# Package options
ggthemr('fresh')  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.2)
options(knitr.kable.NA = '')
```

# The Business Context
A Taiwan-based credit card issuer wants to better predict the likelihood of default for its customers, as well as identify the key drivers that determine this likelihood. This would inform the issuer's decisions on who to give a credit card to and what credit limit to provide. It would also help the issuer have a better understanding of their current and potential customers, which would inform their future strategy, including their planning of offering targeted credit products to their customers. 


<!--
```{r echo=FALSE, message=FALSE}
# Please ENTER the filename
datafile_name = "~/00_INSEAD/03_P3/DSfB/Projects/Nomad2018/train.csv"
ProjectData <- read.csv(datafile_name)
# We turn the data into data.matrix class so that we can easier manipulate it
ProjectData <- data.matrix(ProjectData)
str(ProjectData)

# Please ENTER the dependent variable (class).
# Please use numbers, not column names. E.g., 82 uses the 82nd column as the dependent variable.
# You need to make sure that dependent variable takes only two values: 0 and 1.
dependent_variable = 14

# Please ENTER the attributes to use as independent variables. 
# Please use numbers, not column names. E.g., c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8.
independent_variables = c(2:12) # use all the available attributes

dependent_variable = unique(sapply(dependent_variable,function(i) min(ncol(ProjectData), max(i,1))))
independent_variables = unique(sapply(independent_variables,function(i) min(ncol(ProjectData), max(i,1))))

if (length(unique(ProjectData[,dependent_variable])) !=2){
  cat("\n*****\n BE CAREFUL, THE DEPENDENT VARIABLE TAKES MORE THAN 2 VALUES")
  cat("\nSplitting it around its median...\n*****\n ")
  new_dependent = ProjectData[,dependent_variable] >= median(ProjectData[,dependent_variable])
  ProjectData[,dependent_variable] <- 1*new_dependent
}

# Please ENTER the probability threshold above which an observation is predicted as class 1:
Probability_Threshold = 0.5 # between 0 and 1

# Please ENTER the percentage of data used for estimation
estimation_data_percent = 80
validation_data_percent = 10
test_data_percent = 100-estimation_data_percent-validation_data_percent

# Please ENTER 1 if you want to randomly split the data in estimation and validation/test
random_sampling = 1

# Tree parameter
# Please ENTER the tree (CART) complexity control cp (e.g. 0.0001 to 0.02, depending on the data)
CART_cp = 0.0025
CART_control = rpart.control(cp = CART_cp)

# Please ENTER the words for the business interpretation of class 1 and class 0:
class_1_interpretation = "default"
class_0_interpretation = "no default"

# Please ENTER the profit/cost values for correctly classified and misclassified data:
actual_1_predict_1 = 0
actual_1_predict_0 = -100000
actual_0_predict_1 = 0
actual_0_predict_0 = 20000

Profit_Matrix = matrix(c(actual_1_predict_1, actual_0_predict_1, actual_1_predict_0, actual_0_predict_0), ncol=2)
colnames(Profit_Matrix) <- c(paste("Predict 1 (", class_1_interpretation, ")", sep = ""), paste("Predict 0 (", class_0_interpretation, ")", sep = ""))
rownames(Profit_Matrix) <- c(paste("Actual 1 (", class_1_interpretation, ")", sep = ""), paste("Actual 0 (", class_0_interpretation, ")", sep = ""))

# Please ENTER the maximum number of observations to show in the report and slides 
# (DEFAULT is 50. If the number is large the report and slides may not be generated - very slow or will crash!!)
max_data_report = 10 
```
-->

# The Business Context
A Taiwan-based credit card issuer wants to better predict the likelihood of default for its customers, as well as identify the key drivers that determine this likelihood. This would inform the issuer's decisions on who to give a credit card to and what credit limit to provide. It would also help the issuer have a better understanding of their current and potential customers, which would inform their future strategy, including their planning of offering targeted credit products to their customers. 



<hr>\clearpage

# The Data