---
title: "Data Science for Business Project: Nomad2018 Predicting Transparent Conductors"
author: "Broersen Doortje, Prickaerts Charles, Van Rijn Rens, Feutray Charles, Ammann David"
output:
  html_document:
    css: ../AnalyticsStyles/default.css
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    includes:
      in_header: ../AnalyticsStyles/default.sty
always_allow_html: yes
---

```{r echo=FALSE, message=FALSE}
make_pdf_file = 0 # SET THIS TO 1 IF WE COMPILE PDF FILE, 0 OTHERWISE (FOR HTML)

source("../AnalyticsLibraries/library.R")
source("../AnalyticsLibraries/heatmapOutput.R")

# Package options
ggthemr('fresh')  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.2)
options(knitr.kable.NA = '')
# install.packages("pacman") / in case you do not have this program yet
if("pacman" %in% rownames(installed.packages()) == FALSE) {install.packages("pacman")} 
pacman::p_load("caret","ROCR","lift","glmnet","MASS","e1071") 
#Other packages
install.packages("hexbin")
library(hexbin)
library(ggplot2)


```

# Project Background
Transparent conductors are critical to improve the generation and utilization of energy. Better use of energy is crucial to mitigate environmental challenges imposed by a growing demand of energy. Big limitation in optimization of conductors is that it is hard to predict which potential materials are most promising in terms of bandwith energy and formation energy. 

# Project objectives
The team will participate in the Kaggle competition “Nomad2018 Predicting Transparent Conductors” which will end on February 15, 2018.
* Predict the key properties of novel transparent semiconductors
* Facilitate the discovery of new transparent conductors by predicting bandwith energy and formation energy of promising materials, using other known characteristics of materials
* Compete and achieve top 50% rank in the Kaggle competition 


```{r echo=FALSE, message=FALSE}
datafile_name_train = "../Data Heros Project/Data/train.csv"
datafile_name_submission = "../Data Heros Project/Data/submission.csv"
materials.data.train<-read.csv(datafile_name_train, header=TRUE, sep=",")
materials.data.submission<-read.csv(datafile_name_submission, header=TRUE, sep=",")

# Please ENTER the maximum number of observations to show in the report and slides 
# (DEFAULT is 50. If the number is large the report and slides may not be generated - very slow or will crash!!)
max_data_report = 10 

# Please ENTER the dependent variable (class).
# Please use numbers, not column names. E.g., 82 uses the 82nd column as the dependent variable.
# You need to make sure that dependent variable takes only two values: 0 and 1.
dependent_variable = c(13:14)

# Please ENTER the attributes to use as independent variables. 
# Please use numbers, not column names. E.g., c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8.
independent_variables = c(1:12) # use all the available attributes

dependent_variable = unique(sapply(dependent_variable,function(i) min(ncol(materials.data.train), max(i,1))))
independent_variables = unique(sapply(independent_variables,function(i) min(ncol(materials.data.train), max(i,1))))
```

<hr>\clearpage

# The Data
Data Source: https://www.kaggle.com/c/nomad2018-predict-transparent-conductors/data
The data was provided by the Kaggle challenge provider, including the following data description.

High-quality data are provided for 3,000 materials that show promise as transparent conductors. The following information has been included:

Name                       | Description
:--------------------------|:--------------------------------------------------------------------
ID                         | ID of each combination
Spacegroup                 | A label identifying the symmetry of the materia
Number_of_total_number     | Total number of Al, Ga, In and O atoms in the unit cell 
Percent_atom_al            | Percentage of Al atoms in the structure
Percent_atom_ga            | Percentage of Ga atoms in the structure
Percent_atom_in            | Percentage of In atoms in the structure
Lattice_vector_1_ang       | Lattice vectors lv1 (which are lengths given in units of angstroms (10−1010−10 meters) 
Lattice_vector_2_ang       | Lattice vectors lv2 (which are lengths given in units of angstroms (10−1010−10 meters)
Lattice_vector_3_ang       | Lattice vectors lv3 (which are lengths given in units of angstroms (10−1010−10 meters)
Lattice_angle_alpha_degree | Lattice angle α (angles in degrees between 0° and 360°)
Lattice_angle_beta_degree  | Lattice angle β (angles in degrees between 0° and 360°)
Lattice_angle_gamma_degree | Lattice angle γ (angles in degrees between 0° and 360°)
Formation_energy_ev_natom  | Formation energy (an important indicator of the stability of a material)
Bandgap_energy_ev          | Bandgap energy (an important property for optoelectronic applications)

A domain expert will understand the physical meaning of the above information but those with a data mining background may simply use the data as input for their models.

Good data analytics start with good contextual knowledge as well as a simple statistical and visual exploration of the data. Let's look into the data for a few material combinations.
This is how the first 4 out of the total of `r nrow(materials.data.train)` rows look like (transposed, for convenience):

```{r echo=FALSE, message=FALSE, prompt=FALSE, results='asis'}
knitr::kable({
  df <- t(head(materials.data.train,4))
  colnames(df) <- sprintf("%02d", 1:ncol(df))
  df
})
```

#Data overview for train data
The statistics of our train data looks like this:

```{r echo=FALSE}
knitr::kable(round(my_summary(materials.data.train),2))

par(mfrow=c(2,2))
hist(materials.data.train$formation_energy_ev_natom)
hist(materials.data.train$bandgap_energy_ev)
plot.default(materials.data.train$bandgap_energy_ev ~ materials.data.train$formation_energy_ev_natom)
hist(materials.data.train$spacegroup)
par(mfrow=c(2,2))
hist(materials.data.train$number_of_total_atoms)
hist(materials.data.train$percent_atom_al)
hist(materials.data.train$percent_atom_ga)
hist(materials.data.train$percent_atom_in)
par(mfrow=c(2,2))
plot.default(materials.data.train$bandgap_energy_ev ~ materials.data.train$number_of_total_atoms)
plot.default(materials.data.train$formation_energy_ev_natom ~ materials.data.train$number_of_total_atoms)
plot.default(materials.data.train$percent_atom_al ~ materials.data.train$percent_atom_ga)
plot.default(materials.data.train$percent_atom_in ~ materials.data.train$percent_atom_ga)
```

#Data overview for submission data
The statistics of our submission data looks like this:

```{r echo=FALSE}
knitr::kable(round(my_summary(materials.data.submission),2))

par(mfrow=c(2,2))
hist(materials.data.submission$spacegroup)
hist(materials.data.submission$number_of_total_atoms)
plot.default(materials.data.submission$percent_atom_al ~ materials.data.submission$percent_atom_ga)
plot.default(materials.data.submission$percent_atom_in ~ materials.data.submission$percent_atom_ga)
par(mfrow=c(2,2))
hist(materials.data.submission$percent_atom_al)
hist(materials.data.submission$percent_atom_ga)
hist(materials.data.submission$percent_atom_in)
plot.default(materials.data.submission$spacegroup)
```

#First iteration of regression models 

##Variable reclassification and creationg of training/validation/submissioning sets 

```{r echo=FALSE}
knitr::kable(round(my_summary(materials.data.train),2))

#Summary info of variables 
str(materials.data.train)

## Fixing incorrectly classified data types ##
materials.data.train$spacegroup <- as.factor(materials.data.train$spacegroup)

#Summary info of variables 
str(materials.data.train)

# Creating test
set.seed(500) #set a random number generation seed to ensure that the split is the same everytime
inTrain <- createDataPartition(y = materials.data.train$formation_energy_ev_natom,p = 1918/2400, list = FALSE)
training <- materials.data.train[ inTrain,]
testing <- materials.data.train[ -inTrain,]

```

##Multiple linear regression model (lm) on the training data [Formation]

```{r echo=FALSE}
knitr::kable(round(my_summary(materials.data.train),2))

#Run model and plot results [formation]
fit.f<-lm(formation_energy_ev_natom~.-bandgap_energy_ev, data=training)
predicted.formation<-predict(fit.f,testing) 
plot(predicted.formation ~ testing$formation_energy_ev_natom)

#Run model and plot results [formation]
fit.b<-lm(bandgap_energy_ev~.-formation_energy_ev_natom, data=training)
predicted.bandgap<-predict(fit.b,testing) 
plot(predicted.bandgap ~ testing$bandgap_energy_ev)

#Calculate and show percentage error
percent.errors.f <- sqrt(mean((log(predicted.formation+1) -log(testing$formation_energy_ev_natom+1))^2))
percent.errors.b <- sqrt(mean((log(predicted.bandgap+1) -log(testing$bandgap_energy_ev+1))^2))
mean.score<- (percent.errors.f+percent.errors.b) /2
paste0("score is ",mean.score)

```

##Multiple log regression model on the training data [Formation+Bandgap]

```{r echo=FALSE}
knitr::kable(round(my_summary(materials.data.train),2))

#Run model and plot results [formation]
fit.f.<-glm(log(formation_energy_ev_natom+1)~.-bandgap_energy_ev, data=training, family=gaussian(link = "identity"))
predicted.formation<-exp(predict(fit.f,testing) )-1
plot(predicted.formation ~ testing$formation_energy_ev_natom)

#Run model and plot results [formation]
fit.b<-glm(log(bandgap_energy_ev+1)~.-formation_energy_ev_natom, data=training, family=gaussian(link = "identity"))
predicted.bandgap<-exp(predict(fit.b,testing))-1
plot(predicted.bandgap ~ testing$bandgap_energy_ev)

#Calculate and show percentage error
percent.errors.f <- sqrt(mean((log(predicted.formation+1) -log(testing$formation_energy_ev_natom+1))^2))
percent.errors.b <- sqrt(mean((log(predicted.bandgap+1) -log(testing$bandgap_energy_ev+1))^2))
mean.score<- (percent.errors.f+percent.errors.b) /2
paste0("score is ",mean.score)

```

##Adding Stepwise 

``` {r echo=FALSE}
knitr::kable(round(my_summary(materials.data.train),2))

#Run model and plot results [formation]
fit.f<-lm(formation_energy_ev_natom~.-bandgap_energy_ev, data=training)
fit.f.AIC<-stepAIC(fit.f,direction = c("both"),trace = 1)
predicted.formation<-predict(fit.f.AIC,testing) 
plot(predicted.formation ~ testing$formation_energy_ev_natom)

#Run model and plot results [formation]
fit.b<-lm(bandgap_energy_ev~.-formation_energy_ev_natom, data=training)
fit.b.AIC<-stepAIC(fit.b,direction = c("both"),trace = 1)
predicted.bandgap<-predict(fit.b.AIC,testing) 
plot(predicted.bandgap ~ testing$bandgap_energy_ev)

#Calculate and show percentage error
percent.errors.f <- sqrt(mean((log(predicted.formation+1) -log(testing$formation_energy_ev_natom+1))^2))
percent.errors.b <- sqrt(mean((log(predicted.bandgap+1) -log(testing$bandgap_energy_ev+1))^2))
mean.score<- (percent.errors.f+percent.errors.b) /2
paste0("score is ",mean.score)

```

##Including dimension reduction [Not working]

```{r echo=FALSE}
knitr::kable(round(my_summary(materials.data.train),2))

#All data is confirmed to be metric

#Data scaling 
training.scaled=apply(training,2, function(r){ 
  if (sd(r)!=0) {  
    res=(r-mean(r))/sd(r)  
    } else {  
      res=0*r; res 
      } 
  }) 

#Choose the number of factors
Package: psych 
Package: FactoMineR 

UnRotated_Results<-principal(training, nfactors=ncol(training), rotate="none”, score=TRUE) 

Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE) 
Variance_Explained_Table<-Variance_Explained_Table_results$eig 

``` 